A framework for embedding entities in a textual narrative: a case study on Les Misérables

### 2022-10-12 article/paper1.tex
write article

%%
%% The "title" command^
\title{Characterizing narrative entities with embeddings : a case study of character relationships in Les Misérables}

@@

When a single (or a few) \emph{literary work} is analyzed, a common practice is to study \emph{narrative entities} (characters, event, location, etc.) used by the author in her/his book \cite{Schmid2010}. Researchers are frequently interested in characterizing them and seeing how they are articulated to one another in the story. Various computational tools can help them in this tasks, to name a few, Named Entity Recognition tools \cite{Agarwal2013, Chaturvedi2017, Li2022}, Automatic Character Networks Extraction \cite{Labatut2019}, Sentiment Analysis and Topic Modeling \cite{Min2019}, Textometry \cite{Novakova2019}, and Word Embeddings \cite{Grayson2016, Heuser2017, Kerr2017}. All these methods have been used in order to explicitly show hidden structures constructed by the author in her/his work. It permits to find regularities and patterns, and can help to categorize particular narrative constructions, writing styles, or genres. These kind of methods can be great complement to classical analyses of literary works as they allow to efficiently summarized information which is otherwise quite diffuse.

In this article, we propose a general framework in order to automatically characterize various narrative entities in a literary work. The entire framework is exposed starting from a wide perspective, which is how to organize the textual data, and is narrowed down to a specific use, the study of character relationships in \emph{Les Misérables}, by Victor Hugo. Along this presentation, various choices are made to highlight a particular usage of this framework, but theses choices should be viewed as suggestion rather than rules: the real strength of this framework is its flexibility and the direction taken in this article is oriented for a defined task. To be more specific, we will show how to use \emph{embeddings} in order to locate \emph{character relationships} alongside the vocabulary. An association measure can then be constructed between these words and the relationships, which can help a practitioner to characterize the latters. Four variations of this method are proposed, and are tested on \emph{Les Misérables}.

@@

\begin{table}[h]
	\scriptsize
	\begin{tabular}{|c||c|c|c|c|}
		\hline
		& Cosette & Thénardier & Valjean & ... \\
		\hline
		\hline 
		$u_{101}$ & 1 & 1 & 0 & ... \\
		\hline
		$u_{102}$ & 1 & 1 & 0 & ... \\
		\hline
		$u_{103}$ & 1 & 1 & 0 & ... \\
		\hline
		$u_{104}$ & 1 & 0 & 1 & ... \\
		\hline
	\end{tabular}
	\begin{tabular}{|c||c|c|c|c|}
		\hline
		& ... & Cosette-Thénardier & Cosette-Valjean & ... \\
		\hline
		\hline 
		$u_{101}$  	& ... & 1 & 0 &  ... \\
		\hline
		$u_{102}$ 	& ... & 1 & 0 & ... \\
		\hline
		$u_{103}$  	& ... & 1 & 0 & ... \\
		\hline
		$u_{104}$ 	& ... & 0 & 1 & ... \\
		\hline
	\end{tabular}
	\label{occ_table}
	\caption{A snippet of the unit-entity table $\mathbf{E}$ extracted from \emph{les Misérables}. Rows are chapters, columns are characters (top) and character-pairs (bottom), and cell $e_{ij}$ denotes if $j$ appear in chapter $i$.}
\end{table}

This data organization already gives an orientation to subsequent analyses, and should be kept in mind of the practitioner. Textual unit are now considered as \emph{individuals} (in the statistical terminology), defined by their \emph{variables} contained in the different columns of both tables. Moreover, subsequent analyses are oriented in searching how the unit-entity table $\mathbf{E}$ has an influence over the unit-word table $\mathbf{N}$, i.e. searching which words are over-represented or under-represented knowing entities in the unit. While an authors uses characters in order to build her/his narrative, we, to a certain extent, work backward: we are searching how character appearances and interactions in the textual unit act on her/his choice of words. If the extraction method permits it, a practitioner should include all entities which she/he desires to study. Here, for example, the choice to include character-pairs along with characters is motivated by the fact that we are interested in studying character relationships. A character-pair can roughly be seen as an interaction between two characters, and this interaction should be considered as an object of its own: the presence of this interaction in a unit do not result in having a mixture of words used for each character, but rather gives a specific flavor to the unit.

@@

Various methods can be performed on the contingency table $\mathbf{N}$ in order to extract information from it. Here, we make to choice to extract a lower dimensional, numeric representation of each unit, in other words, a \emph{textual unit vector} located in an \emph{embedding space}. \\
In section \ref{entity_embeddings}, these vectors of textual units are used as anchor points in order to also embed entities into the same space. Therefore, it is crucial that an interpretation about the directions or the regions of this embedding space is possible, in order to properly interpret the localization of entity vectors (the relative position of entity vectors among themselves is generally insufficient). For that reason, we focus on embeddings of textual units which also contain \emph{vectors of words}: by examining the positions of entities relatively to word vectors, entities can be characterized. We propose two embeddings verifying this condition: Section \ref{ca_method} describes \emph{Correspondence Analysis (\textbf{CA})} and section \ref{wv_method} focuses on \emph{Pre-trained Word Vectors (\textbf{WV})}.

@@

When building a regression model with multiple explanatory variables, it is possible to include every variables and their \emph{interactions}. By doing so, we suppose that the effect of raising both variables is not the same as raising each variable independently. Regression models seem therefore appropriate to capture specificities of having a particular entity in a textual unit. For example, in the case of character pairs, the presence of a character $a$ will have a effect on the vocabulary of an unit, the presence of another character $b$ will have another effect, and the presence of the pair $\{a, b\}$ yet a different effect. Now, dependent variables in regression models still need to be defined. In fact, we are doing $d$ regressions, with $d$ the number of dimensions of the embedding, and each regression is constructed to predict the $\alpha$-th coordinate of units by using binary variables in the table $\mathbf{E}$. In matrix notation, all regression models can be written as
\begin{align}
\mathbf{X} = \widetilde{\mathbf{E}} \mathbf{B} + \bm{\Sigma}, \label{cent_sol}
\end{align}
where $\mathbf{X} = (x_{i\alpha})$ is the $(n \times d)$ matrix containing unit vectors (on rows), $\widetilde{\mathbf{E}}$ is the matrix $\mathbf{E}$ with an first additional column of $1$ for the intercept, $\mathbf{B} = (\beta_{k\alpha})$ is the $((p+1) \times d)$ matrix containing intercepts and regression coefficients (each regression is contained on a column), and $\mathbf{\Sigma}$ the $(n \times \alpha)$ matrix containing normal errors.

Estimates $\widehat{\mathbf{B}} = (\widehat{\beta}_{k\alpha})$ for the intercept and coefficient regressions are in fact the researched embeddings for entities as well as for the intercept representing the general tone of the studied text. We therefore denote these estimates with $\mathbf{Y} = (y_{k\alpha})$ in the following, with the convention $y_{0\alpha}$ for intercept coefficients.

As the number of entities (i.e. predictors) might be very large, it is a good idea to add a $L^2$ regularization term in the objective function. Moreover, the quadratic error rate should also be weighted by the number of tokens in each unit. Including all this, we find the solution for our character-based vectors $\mathbf{y}_0, \mathbf{y}_1, \ldots, \mathbf{y}_p$ contained in rows of $\mathbf{Y}$ with 
\begin{equation}
\mathbf{Y} = (\widetilde{\mathbf{E}}^\top \textbf{Diag}(\mathbf{f}) \widetilde{\mathbf{E}} + \lambda \mathbf{I}_{(p+1)})^{-1} \widetilde{\mathbf{E}}^\top \textbf{Diag}(\mathbf{f}) \mathbf{X}, \label{reg_sol}
\end{equation}
where $\textbf{Diag}(\mathbf{f})$ is the diagonal matrix containing weights of units $\mathbf{f} = \left( f_i \right)$, $\lambda > 1$ is the regularization coefficient, and $\mathbf{I}_{(p+1)}$ is the identity matrix of size $((p+1) \times (p+1))$.

@@

For the moment, it is not possible to evaluate the presented methods by the use of some measurements, which would allow to test its validity on various corpora. In order to see if the methods give coherent results, we have to carefully scrutinized and compared them with previous knowledge of the studied work. For this reason, and because of method variations and multiplicity of the results (and lack of place), we chose to present here only one case study: \emph{Les Misérables} by Victor Hugo. The choice of this work is motivated by the fact that it is a large corpus, well-known, immensely studied, and containing various colorful characters and characters relationships. It is therefore a solid choice to clearly illustrate the potential of the presented methodology.

@@

We chose to use chapter as textual units. The table $\mathbf{N}$ (Figure \ref{cont_table}) was build by considering words appearing at least $20$ times in the text and resulted in a table of size $365$ chapters $\times$ $1974$ words.

Characters were detected using \emph{Flair}\footnote{\url{https://github.com/flairNLP/flair}.} NER tools \cite{Akbik2018}. In order to unify character and to further refine the detected characters list, we used hand-made lists of character and aliases from NER results. It resulted in the detection of 54 characters. The entities considered in table $\mathbf{E}$ (Figure \ref{occ_table}) are single character (54 entities) and character pairs (547 entities), resulting in a table of size $365 \times 601$. A character or a character pair is considered present iff characters are detected at least 2 times in the chapter.

Note that, in section \ref{diachronic}, we also tested experiments with entities consisting in character and character-pairs as found in each volumes (e.g. Cosette-Valjean in volume one and Cosette-Valjean in volume two are now two different objects), with the addition of volume constants ($V_i=1$ in volume $i$ and $V_i=0$ in other volumes) in order to isolate volume specific vocabulary. This new table $\mathbf{E}_\text{vol}$, containing $1124$ entities, permits to see a diachronic evolution of words describing volumes, characters, and character relationships.

@@

There are two types of methods for unit embeddings, \textbf{CA} (section \ref{ca_method}) and \textbf{WV} (section \ref{wv_method}), as well as two methods for entity embeddings, \textbf{CENT} (section \ref{centroid}) and \textbf{REG} (section \ref{regression}), making a total of 4 possibles ways for obtaining final embeddings.

The \textbf{CA} method is fully automated, and results in vectors in a $364$-dimensional space, while the $\textbf{WV}$ methods is based on pre-trained word vectors using \emph{fastText} \cite{Bojanowski2017} trained on Common Crawl.\footnote{\url{ https://fasttext.cc/docs/en/crawl-vectors.html}.} For French, the number of word vectors $m$ is 2 millions and the dimension of the vector space is $300$.

Note that, in addition to having 2 tables $\mathbf{E}$ and $\mathbf{E}_\text{vol}$, 4 variations for the methods, and a considerable number of entities, results can also be presented in various ways (similarities between entities, associations between entities and words, etc.). Thus, we chose to show here a selection of results for the each method: the 5 most associated words regarding entities (section \ref{words}), the 5 most associated entities regarding a subset of words (section \ref{objects}), and a diachronic study of top associated words for a subset of entities (section \ref{diachronic}). We invite curious readers to consult results for all words and entities, which can be found in our GitHub repository.\footnote{the "results" folder in \url{https://github.com/gguex/char2char_vectors}.}

@@

We can first observe that \textbf{CA} methods seems to summarize entities with a vocabulary closer to the work, while \textbf{WV} methods tend to frequently use words with a wider scope, with notably more verbs. It results in having the \textbf{WV} methods giving a general feeling for the tone used for describing characters and relationships, while the \textbf{CA} methods can depict very specific objects, locations or events associated with these entities. This behavior can be understood by the nature of unit embeddings: in the \textbf{WV} embedding, word vectors are fixed and do not take into account the actual frequencies of words found in the studied corpora. A character can be close to a word appearing only a few times (or none) in the corpus if this word is located near the vocabulary associated with this character, as semantically similar words are located in the same region of space. By contrast, \textbf{CA} will generally takes into account word frequencies along with specificities in order to describe an entity, and very specific words can define particular regions of space.

@@

\subsubsection{The most associated characters-based objects for a subset of words}
\label{objects}

@@

These results are extracted from the transposed table from the last, and display the most associated entities to a selected set of words. They can be found in Table 4. This type of results can be seen as queries, made from a single word, which output the most associated entities in the work related to that query. We chose here to show top entities related to words "aimer", "rue", "justice" and "guerre", as they represent some of the main topics of the book. In this task again, from our point of view, the \textbf{CA-REG} displays the most accurate results: the main love relationship (Cosette-Marius) of the book is the most associated entity for "aimer", several "amis de l'ABC" (a revolutionary group) are most associated with "rue", the cop-suspect relationship (Javert-Valjean) is the top entity for "justice", and military officers or bellicose characters are associated with "guerre". While somewhat inferior with the selected set of queries, \textbf{WV} methods have the advantage to be able to query words outside the scope of the book, as the pre-trained word embedding possess a very large vocabulary.

@@

\subsubsection{A diachronic study of the most associated words for a subset of character-based objects}
\label{diachronic}

@@

These results are obtained from the table $\mathbf{E}_\text{vol}$ where entities are considered different based on the volume. By doing so, it permits to track the evolution of association scores along the book. Additionally to entities, we can also define constant term $V_i$ for each volumes $i$, which absorbs the associated words with each volume. Results for constants and a subset of entities (Valjean, Cosette, Cosette-Valjean) can be found in Table 5. Note that we did not show \textbf{CONT} results in this table, as they are similar to the one found in Table 3 : words are often repeated themselves for different objects and are less convincing.

Here again, we see that associated words for the \textbf{WV} give a general tone to tomes and entities, while \textbf{CA} results are more specific and related to particular events which occurred for characters. As expected, words associated with volume constants give a short overview of each volumes, especially with the \textbf{CA-REG} method (e.g. $V_2$ for the Battle of Waterloo, $V_4$ for the barricade event). Associated words with entities also seems accurate in describing them. Note that Cosette was not detected in volume 3 because she is not explicitly cited (she is often refereed as "the daugther of M. Leblanc"), and this also explains the absence of the Cosette-Valjean pair.

@@

The first important choice in the analysis is how to define the size of units. Other corpora were also tested (e.g. Shakespeare plays) and it seems important to define units with at least a paragraph size (after preprocessing) in order to represent them accurately. Choosing small units might be able to successfully capture word specificities related to a small subset of entities, but their direction becomes almost orthogonal one to another if the size of units is too small. This situation results in taking into account the "noise" in subsequent analysis, typically denoting an overfitting regime with an high variance and low bias. By contrast, large units will result in an underfitting regime, with a low variance and high bias, failing to capture entities specificities, but more robust to particularities in word usages. Having enough units is also important in order to properly locate entities in the embedding space. In order to analyze characters and relationships, we advice the practitioner to use his prior knowledge of the work in order to split the studied narrative as close as possible to "scenes" (as found in theater), which describes a particular event between an almost constant set of characters.

The second choice is to select which entities to study. This choice is of course driven by the studied problematic, but is also limited by the automatic extraction tools available. These entities can be quite various, but must appear frequently in the work in order to be placed correctly in the embedding space. However, it is unadvised to set an entity which is almost always there (e.g. a narrator), as it will already be represented by the origin in the $\mathbf{CENT}$ method or as the constant term in the $\mathbf{REG}$ method. As a rules of thumbs, the number of entities should ideally be lower than the number of textual units. However, even with an exceeding number of entities (like in our case study, where we had 601 and 1124 entities for 365 units), if some entities appear rarely, analyses are still possible. Note that the version of the table $\mathbf{E}$ containing counts of entities rather than presence among each units was also tested in experiments, but gave similar results for the studied corpus.

The choice of using embeddings, where units, entities and words are located, is motivated by the fact that the resulting space permits many types of explorations. As presented in this article, we can extract some of the most (or least) associated words with each entity or rank entities according to a word query, but other types of measurements could also be made. Entities could be placed along a particular axis in the space, define with a set of positive and a set of negative words, in order to highlight a particular contrast (positive-negative, like in sentiment analysis, introvert-extrovert, friend-enemy, etc.). This approach could also be combined with a clustering of the words, or a Topic Modeling methods, thus permitting to define interpretable regions of the embedding space. Relative location of entities could also be used in order to cluster or classify them. All these leads can be explored by subsequent researches.

The difference in the choice between \textbf{CA} and \textbf{WV} embeddings appears quite clearly in the results. \textbf{CA} highlights particular words associated with entities, very specific to the studied work and the narrative events found in it, while \textbf{WV} give a feel of the general tone in the text when these entities are present. This difference is explained by the fact that \textbf{CA} focuses on words appearing within the work, with possibly very different location to semantically similar words, while \textbf{WV} word vectors are positioned regarding their semantic and syntactic similarities. A entity located in the \textbf{WV} space will then be in a semantic or syntactic region, and its characterizing words should all be related. Results show that \textbf{CA} methods generally perform better to quickly interpret entities among the narrative, but might bit limited for some applications. As a matter of fact, the advantage of \textbf{WV} embedding is that its space is absolute, permitting the comparison of results between sequentially studied texts, and also contains a larger vocabulary in its embedding. This last property could be used in order to use a fixed list of relationship attribute (e.g., friend, enemy, family, colleague), which do not necessarily appear in every text in the studied corpus, in order to categorized character relationships.

The choice between the \textbf{CENT} method and the \textbf{REG} method is relatively easy: thanks to its hyperparameter $\lambda$, the \textbf{REG} method can give similar results to the \textbf{CENT} method when $\lambda$ is high (with only a contraction of entity vectors), but also give more "perpendicular" word affinities between entities when $\lambda$ is low. Thus, it is clearly a superior choice in order to give a variety of results. The choice for this hyperparameter $\lambda$ depends on what the practitioner want from his results. If her/his entities are defined such that some of them are completely included into other, such as character and character-pair, and she/he would like to have specificities about the finer grained entity, the $\lambda$ must be set to a low value. By contrast, if he do not mind to have some of her/his entities defined as a mixture of others, he can set $\lambda$ to a high value. However, very low values of $\lambda$ should be avoided if the number of entities is high compared to the number of units, as this will lead to an overfitting of regression coefficients and will result in the association of very rare and specific words with entities.

>>>>

%%
%% The "title" command^
\title{A framework for embedding entities in a textual narrative: a case study on Les Misérables}

@@

When a single (or a few) \emph{literary work} is analyzed, a common practice is to study \emph{narrative entities} (characters, event, location, etc.) used by the author in her/his book \cite{Schmid2010}. Researchers are frequently interested in depicting them and in seeing how they interact with each other in the story. Various computational tools can help them in this tasks, to name a few, Named Entity Recognition tools \cite{Agarwal2013, Chaturvedi2017, Li2022}, Automatic Character Networks Extraction \cite{Labatut2019}, Sentiment Analysis and Topic Modeling \cite{Min2019}, Textometry \cite{Novakova2019}, and Word Embeddings \cite{Grayson2016, Heuser2017, Kerr2017}. All these methods have been used in order to explicitly show hidden structures constructed by the author in her/his work. It permits to find regularities and patterns, and can help to categorize particular narrative constructions, writing styles, or genres. These kind of methods can be great complement to classical analyses of literary works as they allow to efficiently summarized information which is otherwise quite diffuse.

In this article, we propose a general framework in order to automatically characterize various narrative entities in a literary work. The entire framework is exposed starting from a wide perspective, which is how to organize the textual data, and is narrowed down to a specific use, the study of character relationships in \emph{Les Misérables}, by Victor Hugo. Along this presentation, various choices are made to highlight a particular usage of this framework, but theses choices should be viewed as suggestion rather than rules: the real strength of this framework is its flexibility and the direction taken in this article is oriented for a defined task. To be more specific, we will show how to use \emph{embeddings} in order to locate \emph{character relationships} alongside the vocabulary. An association measure can then be constructed between these words and the relationships, which can help a practitioner to depict the latters. Four variations of this method are proposed, and are tested on \emph{Les Misérables}.

@@

\begin{table}[h]
	\scriptsize
	\begin{tabular}{|c||c|c|c|c|}
		\hline
		& Cosette & Thénardier & Valjean & ... \\
		\hline
		\hline 
		$u_{101}$ & 1 & 1 & 0 & ... \\
		\hline
		$u_{102}$ & 1 & 1 & 0 & ... \\
		\hline
		$u_{103}$ & 1 & 1 & 0 & ... \\
		\hline
		$u_{104}$ & 1 & 0 & 1 & ... \\
		\hline
	\end{tabular}
	\begin{tabular}{|c||c|c|c|c|}
		\hline
		& ... & Cosette-Thénardier & Cosette-Valjean & ... \\
		\hline
		\hline 
		$u_{101}$  	& ... & 1 & 0 &  ... \\
		\hline
		$u_{102}$ 	& ... & 1 & 0 & ... \\
		\hline
		$u_{103}$  	& ... & 1 & 0 & ... \\
		\hline
		$u_{104}$ 	& ... & 0 & 1 & ... \\
		\hline
	\end{tabular}
	\label{occ_table}
	\caption{A snippet of the unit-entity table $\mathbf{E}$ extracted from \emph{les Misérables}. Rows are chapters, columns are characters (top) and character pairs (bottom), and cell $e_{ij}$ denotes if $j$ appear in chapter $i$.}
\end{table}

This data organization already gives an orientation to subsequent analyses, and should be kept in mind of the practitioner. Textual unit are now considered as \emph{individuals} (in the statistical terminology), defined by their \emph{variables} contained in the different columns of both tables. Moreover, subsequent analyses are oriented in searching how the unit-entity table $\mathbf{E}$ has an influence over the unit-word table $\mathbf{N}$, i.e. searching which words are over-represented or under-represented knowing entities in the unit. While an authors uses characters in order to build her/his narrative, we, to a certain extent, work backward: we are searching how character appearances and interactions in the textual unit act on her/his choice of words. If the extraction method permits it, a practitioner should include all entities which she/he desires to study. Here, for example, the choice to include character pairs along with characters is motivated by the fact that we are interested in studying character relationships. A character pair can roughly be seen as an interaction between two characters, and this interaction should be considered as an object of its own: the presence of this interaction in a unit do not result in having a mixture of words used for each character, but rather gives a specific flavor to the unit.

@@

Various methods can be performed on the contingency table $\mathbf{N}$ in order to extract information from it. Here, we make to choice to extract a lower dimensional, numeric representation of each unit, in other words, a \emph{textual unit vector} located in an \emph{embedding space}. \\
In section \ref{entity_embeddings}, these vectors of textual units are used as anchor points in order to also embed entities into the same space. Therefore, it is crucial that an interpretation about the directions or the regions of this embedding space is possible, in order to properly interpret the localization of entity vectors (the relative position of entity vectors among themselves is generally insufficient). For that reason, we focus on embeddings of textual units which also contain \emph{vectors of words}: by examining the positions of entities relatively to word vectors, entities can be depicted. We propose two embeddings verifying this condition: Section \ref{ca_method} describes \emph{Correspondence Analysis (\textbf{CA})} and section \ref{wv_method} focuses on \emph{Pre-trained Word Vectors (\textbf{WV})}.

@@

When building a regression model with multiple explanatory variables, it is possible to include every variables and their \emph{interactions}. By doing so, we suppose that the effect of raising both variables is not the same as raising each variable independently. Regression models seem therefore appropriate to capture specificities of having a particular entity in a textual unit. For example, in the case of character pairs, the presence of a character $a$ will have a effect on the vocabulary of an unit, the presence of another character $b$ will have another effect, and the presence of the pair $\{a, b\}$ yet a different effect. Now, dependent variables in regression models still need to be defined. In fact, we are doing $d$ regressions, with $d$ the number of dimensions of the embedding, and each regression is constructed to predict the $\alpha$-th coordinate of units by using binary variables in the table $\mathbf{E}$. In matrix notation, all regression models can be written as
\begin{align}
\mathbf{X} = \widetilde{\mathbf{E}} \mathbf{B} + \bm{\Sigma}, \label{cent_sol}
\end{align}
where $\mathbf{X} = (x_{i\alpha})$ is the $(n \times d)$ matrix containing unit vectors (on rows), $\widetilde{\mathbf{E}}$ is the matrix $\mathbf{E}$ with an first additional column of $1$ for the intercept, $\mathbf{B} = (\beta_{k\alpha})$ is the $((p+1) \times d)$ matrix containing intercepts and regression coefficients (each column corresponds to one regression), and $\mathbf{\Sigma}$ the $(n \times \alpha)$ matrix containing normal errors.

Intercepts and coefficients estimations $\widehat{\mathbf{B}} = (\widehat{\beta}_{k\alpha})$ can be considered as our embeddings for entities as well as for the intercept, which represents the general tone of the studied text. We therefore denote these estimates with $\mathbf{Y} = (y_{k\alpha})$ in the following, with the notation convention $y_{0\alpha}$ for intercept coordinate $\alpha$.

As the number of entities (i.e. predictors) might be very large, it is a good idea to add a $L^2$ regularization term in the objective function. Moreover, the quadratic error rate should also be weighted by the number of tokens in each unit. Including all this, we find the solution for our intercept and entity vectors $\mathbf{y}_0, \mathbf{y}_1, \ldots, \mathbf{y}_p$, contained in the rows of $\mathbf{Y}$, with 
\begin{equation}
\mathbf{Y} = (\widetilde{\mathbf{E}}^\top \textbf{Diag}(\mathbf{f}) \widetilde{\mathbf{E}} + \lambda \mathbf{I}_{(p+1)})^{-1} \widetilde{\mathbf{E}}^\top \textbf{Diag}(\mathbf{f}) \mathbf{X}, \label{reg_sol}
\end{equation}
where $\textbf{Diag}(\mathbf{f})$ is the diagonal matrix containing weights of units $\mathbf{f} = \left( f_i \right)$, $\lambda > 1$ is the regularization coefficient, and $\mathbf{I}_{(p+1)}$ is the identity matrix of size $((p+1) \times (p+1))$.

@@

At the time of writing, it is not possible to evaluate the exposed framework with some kind of metric, which would allow to test its pertinence on various corpora. In order to see if the methods give coherent results, we have to carefully scrutinized and compared them with previous knowledge of the studied work. For this reason, and because of method variations and multiplicity of the results (and lack of place), we chose to present only one case study: the analysis of characters and relationships in \emph{Les Misérables}, by Victor Hugo. The choice of this work is motivated by the fact that it is a large corpus, well-known, immensely studied, and containing various colorful characters and characters relationships. It is therefore a solid choice to clearly illustrate the potential of the exposed framework.

@@

We chose to use chapters as textual units. The table $\mathbf{N}$ (Figure \ref{cont_table}) was build by considering words appearing at least $20$ times in the text and resulted in a table of size $365$ chapters $\times$ $1974$ words.

Characters were detected using \emph{Flair}\footnote{\url{https://github.com/flairNLP/flair}.} NER tools \cite{Akbik2018}. In order to unify characters and to further refine the results, we used hand-made lists of character names and aliases from NER results. It resulted in the detection of 54 characters. The entities considered in table $\mathbf{E}$ (Figure \ref{occ_table}) are composed of 54 single characters and 547 character pairs, resulting in a table of size $365 \times 601$. A character (resp. a pair of characters) is considered present if it is (resp. both are) detected at least 2 times in the chapter.

Note that, in section \ref{diachronic}, we also tested experiments with entities consisting in characters and character pairs as found in each volumes (e.g. Cosette-Valjean in volume one and Cosette-Valjean in volume two are now two different entities), with the addition of volume constants ($V_i=1$ in volume $i$ and $V_i=0$ in other volumes) in order to isolate volume specific vocabulary. This new table $\mathbf{E}_\text{vol}$, containing $1124$ entities, permits to see a diachronic evolution of words associated with volumes, characters, and character relationships.

@@

There are two types of methods for unit embeddings, \textbf{CA} (section \ref{ca_method}) and \textbf{WV} (section \ref{wv_method}), as well as two methods to derive entity embeddings from them, \textbf{CENT} (section \ref{centroid}) and \textbf{REG} (section \ref{regression}), making a total of 4 possibles ways for obtaining entity embeddings.

The \textbf{CA} method do not need any external data, and results in vectors in a $364$-dimensional space, while the $\textbf{WV}$ methods is based on pre-trained word vectors using \emph{fastText} \cite{Bojanowski2017} trained on Common Crawl.\footnote{\url{ https://fasttext.cc/docs/en/crawl-vectors.html}.} For French, the number of word vectors is 2 millions and the dimension of the vector space is $300$.

Note that, in addition to having 2 tables $\mathbf{E}$ and $\mathbf{E}_\text{vol}$, 4 methods, and a considerable number of words and entities, results can also be presented in various ways (similarities between entities, associations between entities and words, etc.). Thus, we chose to show here a selection of results for the each method: the 5 most associated words regarding a subset of entities (section \ref{words}), the 5 most associated entities regarding a subset of words (section \ref{objects}), and a diachronic study of the 5 most associated words for a subset of entities (section \ref{diachronic}). We invite curious readers to consult results for all words and entities, which can be found in our GitHub repository.\footnote{in the "results" folder in \url{https://github.com/gguex/char2char_vectors}.}

@@

We can first observe that \textbf{CA} methods seems to summarize entities with a vocabulary closer to the work, while \textbf{WV} methods tend to frequently use words with a wider scope, with notably more verbs. It results in having the \textbf{WV} methods giving a general feeling for the tone used for describing characters and relationships, while the \textbf{CA} methods can depict very specific objects, locations or events associated with these entities. This behavior can be understood by the nature of unit embeddings: in the \textbf{WV} embedding, word vectors are fixed and do not take into account the actual frequencies of words found in the studied corpora. A character can be close to a word appearing only a few times (or none) in the corpus if this word is located near the vocabulary associated with this character, as semantically similar words are located in the same region of space. By contrast, \textbf{CA} will generally takes into account word frequencies along with specificities in order to describe an entity, and semantically similar words can be located far away from each others.

@@

\subsubsection{The most associated entities for a subset of words}
\label{objects}

@@

These results are extracted from a transposed table, and display the most associated entities to a selected set of words. They can be found in Table 4. This type of results can be seen as queries, made from a single word by a practitioner, which output the most associated entities in the work related to that query. We chose here to show top entities related to words "aimer", "rue", "justice" and "guerre", as they represent some of the main topics of the book. In this task again, from our point of view, the \textbf{CA-REG} displays the most accurate results: the main love relationship (Cosette-Marius) of the book is the most associated entity for "aimer", several "amis de l'ABC" (a revolutionary group) are most associated with "rue", the cop-suspect relationship (Javert-Valjean) is the top entity for "justice", and military officers or bellicose characters are associated with "guerre". While somewhat inferior with the selected set of queries, \textbf{WV} methods have the advantage to be able to query words outside the scope of the book, as the pre-trained word embedding possess a very large vocabulary.

@@

\subsubsection{A diachronic study of the most associated words for a subset of entities}
\label{diachronic}

@@

These results are obtained from the table $\mathbf{E}_\text{vol}$ where entities are considered different based on the volume. By doing so, it permits to track the evolution of association scores along the book. Additionally to entities, we can also define a constant term $V_i$ for each volumes $i$, which absorbs the associated words with each volume. Results for constants and a subset of entities (Valjean, Cosette, Cosette-Valjean) can be found in Table 5. Note that we did not show \textbf{CENT} results in this table, as they are similar to the one found in Table 3: words are often repeated for different entities and are less convincing.

Here again, we see that associated words for the \textbf{WV} give the general tone of tomes and entities, while \textbf{CA} results are more specific and related to particular events which occurred for characters. As expected, words associated with volume constants give a short overview of each volumes, especially with the \textbf{CA-REG} method (e.g. $V_2$ for the Battle of Waterloo, $V_4$ for the barricade event). Associated words with entities also seems accurate in describing them. Note that Cosette was not detected in volume 3 because she is not explicitly cited (she is often refereed as "the daugther of M. Leblanc"), and this also explains the absence of the Cosette-Valjean pair.

@@

The first important choice in the analysis is how to define the size of units. Other corpora were also tested (e.g. Shakespeare plays) and it seems important to define units with at least a paragraph size (after preprocessing) in order to represent them accurately. Choosing small units might be able to successfully capture word specificities related to a small subset of entities, but unit vectors become almost orthogonal one to another if the size of units is too small. This situation results in an overfitting regime with an high variance and low bias, i.e. units positions can be very different with the difference of only a few rare words. By contrast, large units will result in an underfitting regime, with a low variance and high bias, failing to capture entities specificities, but more robust to particularities in word usages. Having enough units is also important in order to properly locate entities in the embedding space. In order to analyze characters and relationships, we advice the practitioner to use his prior knowledge of the work in order to split the studied narrative as close as possible to "scenes" (as found in theater), which describes a particular event between an almost constant set of characters.

The second choice is to select which entities to study. This choice is of course driven by the problematic, but is also limited by the automatic extraction tools available. These entities can be quite various, but must appear frequently in the work in order to be placed correctly in the embedding space. However, it is unadvised to set an entity which is almost always there (e.g. a narrator), as it will already be represented by the origin in the $\mathbf{CENT}$ method or as the constant term in the $\mathbf{REG}$ method. As a rule of thumbs, the number of entities should ideally be lower than the number of textual units. However, even with an exceeding number of entities (like in our case study, where we had 601 and 1124 entities for 365 units), if some entities appear rarely, analyses are still possible. Note that the version of the table $\mathbf{E}$ containing counts of entities rather than presence among each units was also tested in experiments, but gave similar results for the studied corpus.

The choice of using embeddings, where units, entities and words are located, is motivated by the fact that the resulting space permits many types of explorations. As presented in this article, we can extract some of the most (or least) associated words with each entity or rank entities according to a word query, but other types of measurements could also be made. Entities could be placed along a particular axis in the space, define with a set of positive and a set of negative words, in order to highlight a particular contrast (positive-negative, like in sentiment analysis, introvert-extrovert, friend-enemy, etc.). This approach could also be combined with a clustering of the words, or a Topic Modeling method, thus permitting to further refine the different regions in the embedding space. Relative location of entities could also be used in order to cluster or classify them. All these leads can be explored in future researches.

The difference in the choice between \textbf{CA} and \textbf{WV} embeddings appears quite clearly in the results. \textbf{CA} highlights particular words associated with entities, very specific to the studied work and the narrative events found in it, while \textbf{WV} gives a general feeling of the tone of the text when these entities are present. This difference is explained by the fact that \textbf{CA} focuses on words appearing within the work, with possibly very different locations to semantically similar words, while \textbf{WV} word vectors are positioned regarding their semantic and syntactic similarities. A entity located in the \textbf{WV} space will then be in a semantic or syntactic region, and its characterizing words should all be related. Results show that \textbf{CA} methods generally perform better to quickly interpret entities among the narrative, but might bit limited for some applications. As a matter of fact, the advantage of \textbf{WV} embeddings is that its space is absolute, permitting the comparison of results between sequentially studied texts, and also contains a larger vocabulary in its embedding. This last property could be used in order to use a fixed list of relationship attributes (e.g., friend, enemy, family, colleague), which do not necessarily appear in every text in the studied corpus, in order to categorized character relationships.

The choice between the \textbf{CENT} method and the \textbf{REG} method is relatively easy: thanks to its hyperparameter $\lambda$, the \textbf{REG} method can give similar results than the \textbf{CENT} method when $\lambda$ is high (with only a contraction of entity vectors), but also gives more "perpendicular" sets of words describing entities when $\lambda$ is low. Thus, it is clearly a superior choice in order to give a variety of results. The choice for this hyperparameter $\lambda$ depends on what the practitioner desire. If her/his entities are defined such that some of them are completely included into others, such as character and character pair, and she/he would like to have specificities about the finer grained entity, the $\lambda$ must be set to a low value. By contrast, if he do not mind in having some of her/his entities described like a mixture of others, she/he can set $\lambda$ to a high value. However, very low values of $\lambda$ should be avoided if the number of entities is high compared to the number of units, as this will lead to an overfitting of regression coefficients and result in the association of very rare and specific words with entities.

Finally, the biggest weakness of this framework yet is the difficulty to validate its pertinence. Several other case studies, with results carefully scrutinized regarding prior knowledge, should be undertaken in order to see if results are trustworthy, but this type of experiments are expansive both in time and human resources. Another idea could be to use annotated corpora such as described in \cite{Massey2015}, where human annotators classified character relationships in various categories. For example, we could see if the presented method can actually retrieve these categories by assigning relationships to the category-word with the highest association score. Such experiments are promising, but it requires an efficient automatic entity tagger, in order to detect and especially unify characters in a large quantity of documents, and unfortunately, this tool do not exist yet. Nevertheless, first case studies gave promising results for this framework, and its flexibility could lead to various applications.
